Supplemental Information for: The Magnetorotational Instability Prefers Three Dimensions

### 2019-08-28 paper/supplemental.tex
tweaks to supp.

>>>>

\section{1st-order formulation of the dissipative equations}
\label{sec:formulation}

We formulate the system in terms of the 10 problem \texttt{variables} 
\Beq
X \ = \ \left[\,p,v_{x},v_{y},v_{z},\omega_{y},\omega_{z},b_{x},b_{y},b_{z},j_{xx}\,\right]^{\top} \nonumber
\Eeq
We use sting-based \texttt{substitutions} to expedite the input of the full the system of equations: 
\Beq
``\omega_{x}" &=& ``\pd{y} v_{z} - \pd{z} v_{y}" \nonumber \\
``j_{x}" &=& ``\pd{y} b_{z} - \pd{z} b_{y}" \nonumber \\
``j_{y}" &=& ``\pd{z} b_{x} - \pd{x} b_{z}" \nonumber \\
``j_{z}" &=& ``\pd{x} b_{y} - \pd{y} b_{x}" \nonumber 
\Eeq
The expression for $\omega_{x},\,j_{x},\,j_{y},\,j_{z}$ are not legitimate variables that \textit{Dedalus} solves for. They are only short-hand aliases that mean literally the same thing as the expression on the right-hand side of the definition, which are defined strictly in terms of legitimate problem variables.

We also define the following operator substitutions (with general argument $A$)
\Beq
``\pd{y} A" &=& ``i k_{y} A" \nonumber \\
``\pd{z} A" &=& ``i k_{z} A" \nonumber \\
``D_{t} A" &=& ``\sigma q +S x \pd{y} A " \nonumber \\
``\Delta A" &=& ``\pd{y}^{2} A + \pd{z}^{2}A"
\Eeq
where $\sigma$ is the problem \texttt{eigenvalue}, and $k_{y},\,k_{z}$ are problem \texttt{parameters}, along with $f,\,S,\,B_{0},\, \nu,\, \eta$.

The system of \texttt{equations} is:
\Beq\label{u-eq}
\pd{x} v_{x} + \pd{y} v_{y} + \pd{z} v_{z}  &=& 0\\
D_{t} v_{x}  - f v_{y} + \pd{x} p - B_{0} \pd{z} b_{x}  + \nu ( \pd{y} \omega_{z} - \pd{z} \omega_{y}) &=& 0\\
D_{t} v_{y}  + (f+S) v_{x} + \pd{y} p - B_{0} \pd{z} b_{y} + \nu ( \pd{z} \omega_{x} - \pd{x} \omega_{z})&=& 0\label{v-eq} \\
D_{t} v_{z} + \pd{z} p - B_{0} \pd{z} b_{z} + \nu ( \pd{x} \omega_{y} - \pd{y} \omega_{x})&=& 0 \\
\omega_{y} - \pd{z} v_{x} + \pd{x} v_{z} &=& 0 \\
\omega_{z} - \pd{x} v_{y} + \pd{y} v_{x} &=& 0 \\
\pd{x} b_{x} + \pd{y} b_{y} + \pd{z} b_{z}   &=& 0\\
D_{t} b_{x} - B_{0} \pd{z} v_{x} + \eta ( \pd{y} j_{z} - \pd{z} j_{y})&=& 0\\
D_{t} j_{x}  - \eta( \pd{x}j_{xx} +  \Delta j_{x} )  - B_{0} \pd{z} \omega_{x} + S \pd{z} b_{x} &=& 0\\
j_{xx} - \pd{x} j_{x} &=& 0
\Eeq
We can see there are the same number of equations as variables, each equations has only 1st-order derivatives in $x$, and every variable has at least one $\pd{x}$. It appears (for example) that there are no $\pd{x}$ terms in the $b_{x}$ equation, but this due to the substitutions, 
\Beq
\pd{y} j_{z} - \pd{z} j_{y} =  \pd{y}(\pd{x} b_{y} - \pd{y} b_{x}) - \pd{z}(\pd{z} b_{x} - \pd{x} b_{z})  =  \pd{y}\pd{x}b_{y} + \pd{z}\pd{x}b_{z} - (\pd{y}^{2} + \pd{z}^{2})b_{x}.
\Eeq
Also,
\Beq
\pd{x}j_{x} \ = \ \pd{y} \pd{x} b_{z} - \pd{z} \pd{x}b_{y}.
\Eeq
The entire system is formulated and solved as a generalized eigenvalue problem with non-constant coefficients 
\Beq
\sigma \, M \cdot X + L_{0}(x) \cdot X + L_{1}\cdot \pd{x} X \ = \ 0.
\Eeq
The problem is ``generalized" because $M$ is not full rank. The system is 10th order because $L_{1}$ is full rank. We therefore impose the 10 \texttt{boundary conditions} 
\Beq
v_{x} \ = \ \omega_{y} \ = \ \omega_{z} \ = \ b_{x} \ = \ j_{xx} \ = \ 0 \quad \text{at} \quad x \ = \ \pm d/2
\Eeq
We could choose any other set of boundary conditions constructed out of a linearly independent combinations of the legitimate problem variables. We choose impenetrable stress-free and perfectly conducting because it conforms with the purely impenetrable conditions for an ideal calculation, thus minimizing the arbitrary influence of boundary layers.
